---
title: "Data_replication!"
format: html
Author: Hector!
---

## Introduction

Peripheral nerve injuries can severely impair motor function, and while nerve transfer surgery offers a promising path to recovery, predicting individual outcomes remains a significant clinical challenge. Zhu et al. developed the Fully Implantable Multisite Optogenetic Stimulation System (FIMOSS), a wireless, reprogrammable device designed for long-term evaluation of peripheral nerve plexus function at sub-nerve resolution. Using transgenic Thy1-ChR2-EYFP mice, the system delivered targeted optical stimulation to individual nerve fascicles and recorded both electrophysiological and motor responses. The authors then applied FIMOSS to a nerve transfer mouse model following traumatic brain injury, tracking innervation patterns across multiple muscles over 12 weeks post-surgery. Their findings demonstrate that FIMOSS can provide refined, longitudinal assessment of nerve function, with potential applications in personalized diagnosis and treatment planning following nerve injuries or surgeries.
\

## Data and Reanalyses

the data used here was pulled from the supplementary materials of the original paper - they gave raw numbers for several figures 🎉. this report focuses on fig 3, which covers thermal safety, optical output, and stimulation timing accuracy of the FIMOSS implant.

the 4 analyses replicated here:

1. **Visualization (Fig. 3e)** — line plot w/ shaded SEM ribbons, in vitro temp over 7200s at 4 voltages (n=3 per voltage)
2. **Descriptive stats (Fig. 3g)** — mean ± SEM in vivo temp across 4 devices at 5 timepoints over 3600s
3. **Inferential stats (Fig. 3e)** — one-way ANOVA + tukey HSD on steady-state temps across voltage conditions
4. **Visualization (Fig. 3p)** — dual axis plot, temporal accuracy across pulse widths 10-200ms (n=4, mean ± SEM)

## Variable Dictionary

| variable | description |
|---|---|
| `time` | timepoint in seconds |
| `temp` | temperature in °C |
| `voltage` | input voltage condition (3.3V–4.2V) |
| `device` | individual implant device (1–4) |
| `pulse_width` | LED pulse width in ms |
| `mean` | mean across replicates |
| `sem` | standard error of the mean |
| `delta` | temp change from baseline |
| `pct_summary` | time deviation as % |
| `ms_summary` | time deviation in ms |

## loading packages & data 

```{r}
library(tidyverse)
library(readxl)
library(fmsb)

# download data file from GitHub to a temp path
data_url  <- "https://raw.githubusercontent.com/hjleon19/data-analysis-replication/main/data_replication_data.xlsx"
data_path <- tempfile(fileext = ".xlsx")
download.file(data_url, destfile = data_path, mode = "wb")

# main data file
fig3_raw <- read_excel(data_path, sheet = "Fig.3")

# peek at raw
head(fig3_raw)
```

---

## Descriptive Statistics: Fig 3g

###in vivo temp across 4 devices over time - pretty simple stuff
###reading in the specific range for 3g, its buried in the sheet

```{r}
# read in fig3g - in vivo temp data
fig3g <- read_excel(data_path,
                    sheet = "Fig.3",
                    range = "A41:E46",
                    col_names = TRUE)

# rename cols
colnames(fig3g) <- c("time", "Device1", "Device2", "Device3", "Device4")

# peek
head(fig3g)
```

###pivot to long so we can group by device/time easily

```{r}
# long format
fig3g_long <- fig3g |>
  pivot_longer(-time, names_to = "device", values_to = "temp")

head(fig3g_long)
```

###now get mean ± sem per timepoint across the 4 devices

```{r}
# descriptive stats per timepoint
fig3g_summary <- fig3g_long |>
  group_by(time) |>
  summarise(
    n    = n(),
    mean = mean(temp),
    sd   = sd(temp),
    sem  = sd / sqrt(n),
    min  = min(temp),
    max  = max(temp),
    .groups = "drop"
  )

# print the table so we can compare to paper
knitr::kable(fig3g_summary, digits = 3,
             caption = "mean ± SEM in vivo temp per timepoint (Fig 3g)")
```

###also calc delta from baseline per device - shows how much warming happened

```{r}
# temp change from t=0 per device
fig3g_delta <- fig3g_long |>
  group_by(device) |>
  mutate(baseline = temp[time == 0],
         delta    = temp - baseline) |>
  ungroup()

# summarise delta across devices
delta_summary <- fig3g_delta |>
  group_by(time) |>
  summarise(
    mean_delta = mean(delta),
    sem_delta  = sd(delta) / sqrt(n()),
    .groups = "drop"
  )

knitr::kable(delta_summary, digits = 3,
             caption = "mean temp change from baseline per timepoint")
```

###plot - individual device points + dashed mean line w error bars, matches fig 3g style

```{r}
ggplot() +
  geom_point(data = fig3g_long,
             aes(x = time, y = temp, color = device), size = 3) +
  geom_line(data = fig3g_summary,
            aes(x = time, y = mean), linetype = "dashed", linewidth = 0.8) +
  geom_errorbar(data = fig3g_summary,
                aes(x = time, ymin = mean - sem, ymax = mean + sem),
                width = 100) +
  labs(x     = "Time/s",
       y     = "Maximum temperature/°C",
       title = "In Vivo Implant Temperature Curve",
       color = NULL)
```

across the four implanted devices, mean in vivo temperature remained stable and close to baseline throughout the 3600s recording period, ranging from 33.4°C (± 0.03 SEM) at t=0 to a peak of 34.4°C (± 0.39 SEM) at t=1800s, consistent with the modest temperature increases shown in Figure 3g of the original study. mean temp increase from baseline peaked at t=1800s and t=2700s with an avg rise of 0.975°C (± 0.40 and ± 0.46 SEM respectively), and remained below 1°C across all timepoints, indicating that continuous FIMOSS operation at 3.3V produces only minimal subcutaneous heating in vivo 🌡️

---

## Visualization: Fig 3e

###in vitro temp curves - 4 voltages, 3 replicates each, 7200s total
###need to reshape from wide (one col per replicate) to long

```{r}
# read in fig3e
fig3e <- read_excel(data_path,
                    sheet = "Fig.3",
                    range = "A2:M37",
                    col_names = TRUE)

# rename - 3 reps per voltage condition
colnames(fig3e) <- c("time",
                     "v3.3_1", "v3.3_2", "v3.3_3",
                     "v3.6_1", "v3.6_2", "v3.6_3",
                     "v3.9_1", "v3.9_2", "v3.9_3",
                     "v4.2_1", "v4.2_2", "v4.2_3")

head(fig3e)
```

###pivot long + extract voltage label from col name

```{r}
# long format + voltage label
fig3e_long <- fig3e |>
  pivot_longer(-time, names_to = "col", values_to = "temp") |>
  mutate(voltage = str_extract(col, "v[0-9.]+"),
         voltage = recode(voltage,
                          "v3.3" = "3.3V",
                          "v3.6" = "3.6V",
                          "v3.9" = "3.9V",
                          "v4.2" = "4.2V"))

head(fig3e_long)
```

###summarise - mean + min/max for ribbon

```{r}
# mean, min, max per voltage per timepoint
fig3e_summary <- fig3e_long |>
  group_by(time, voltage) |>
  summarise(mean_temp = mean(temp, na.rm = TRUE),
            min_temp  = min(temp,  na.rm = TRUE),
            max_temp  = max(temp,  na.rm = TRUE),
            sem_temp  = sd(temp, na.rm = TRUE) / sqrt(n()),
            .groups = "drop")

knitr::kable(head(fig3e_summary, 12), digits = 3,
             caption = "sample of fig3e summary - mean temp per voltage per timepoint")
```

###plot - ribbon = min/max range, line = mean. matches fig 3e layout

```{r}
ggplot(fig3e_summary,
       aes(x = time, y = mean_temp, color = voltage, fill = voltage)) +
  geom_ribbon(aes(ymin = min_temp, ymax = max_temp),
              alpha = 0.3, color = NA) +
  geom_line() +
  labs(x     = "Time/s",
       y     = "Maximum temperature/°C",
       title = "In Vitro Implant Temperature Curve",
       color = NULL,
       fill  = NULL)
```

---

## Inferential Statistics: Fig 3e 📊

###anova to test if steady state temps differ across voltages
###paper defines long term as t > 200s so filtering to that

```{r}
# filter to steady state (long term phase only)
fig3e_steady <- fig3e_long |>
  filter(time > 200)

# descriptive summary per voltage first
steady_desc <- fig3e_steady |>
  group_by(voltage) |>
  summarise(
    n    = n(),
    mean = mean(temp, na.rm = TRUE),
    sd   = sd(temp,   na.rm = TRUE),
    sem  = sd / sqrt(n),
    .groups = "drop"
  )

knitr::kable(steady_desc, digits = 3,
             caption = "steady state temp summary per voltage (t > 200s)")
```

###one way anova - is voltage a sig predictor of temp?

```{r}
# anova
anova_result <- aov(temp ~ voltage, data = fig3e_steady)
summary(anova_result)
```

###tukey post hoc - which pairs are diff from each other

```{r}
# tukey HSD
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)
```

a one-way ANOVA revealed a significant effect of input voltage on steady-state implant temperature (p < 0.001). post-hoc Tukey HSD testing confirmed that all four voltage conditions (3.3V, 3.6V, 3.9V, and 4.2V) produced significantly different mean temperatures from one another (all adjusted p < 0.001), with higher voltages consistently producing greater heating, replicating the pattern reported in the original study.

---

## Second Visualization: Fig 3p

###dual axis plot - bars = ms deviation, line = % deviation
###two separate tables in the sheet so need to read both

```{r}
# read pct deviation table
fig3p_pct <- read_excel(data_path,
                        sheet = "Fig.3",
                        range = "A89:G92",
                        col_names = TRUE)

# read ms deviation table
fig3p_ms <- read_excel(data_path,
                       sheet = "Fig.3",
                       range = "A96:G100",
                       col_names = TRUE)

# rename both
colnames(fig3p_pct) <- c("rep", "pw10", "pw20", "pw30", "pw40", "pw50", "pw200")
colnames(fig3p_ms)  <- c("rep", "pw10", "pw20", "pw30", "pw40", "pw50", "pw200")

# peek
head(fig3p_pct)
head(fig3p_ms)
```

###pivot + summarise both tables - same function works for both

```{r}
# summarise helper func
summarise_fig3p <- function(df, measure_name) {
  df |>
    pivot_longer(-rep, names_to = "pulse_width", values_to = "value") |>
    mutate(pulse_width = as.numeric(str_remove(pulse_width, "pw"))) |>
    group_by(pulse_width) |>
    summarise(mean = mean(value),
              sem  = sd(value) / sqrt(n()),
              .groups = "drop") |>
    mutate(measure = measure_name)
}

pct_summary <- summarise_fig3p(fig3p_pct, "pct")
ms_summary  <- summarise_fig3p(fig3p_ms,  "ms")

# print both summary tables for comparison to paper
knitr::kable(pct_summary, digits = 4,
             caption = "time deviation % summary per pulse width")
knitr::kable(ms_summary,  digits = 4,
             caption = "time deviation ms summary per pulse width")
```

###dual axis plot - scale factor maps ms onto pct axis
###scale_factor = 5 bc pct goes to ~1.5 and ms goes to ~0.30

```{r}
scale_factor <- 5

ggplot() +
  # bars = ms (scaled up to sit on pct axis)
  geom_col(data = ms_summary,
           aes(x = factor(pulse_width), y = mean * scale_factor),
           fill = "steelblue", width = 0.6) +
  geom_errorbar(data = ms_summary,
                aes(x = factor(pulse_width),
                    ymin = (mean - sem) * scale_factor,
                    ymax = (mean + sem) * scale_factor),
                width = 0.2) +
  # line = pct deviation
  geom_line(data = pct_summary,
            aes(x = factor(pulse_width), y = mean, group = 1),
            color = "red", linewidth = 1) +
  geom_point(data = pct_summary,
             aes(x = factor(pulse_width), y = mean),
             color = "red", size = 2) +
  geom_errorbar(data = pct_summary,
                aes(x = factor(pulse_width),
                    ymin = mean - sem,
                    ymax = mean + sem),
                color = "red", width = 0.2) +
  # dual axis - right axis divides back by scale factor to show real ms vals
  scale_y_continuous(
    name = "Time Deviation/%",
    sec.axis = sec_axis(~ . / scale_factor, name = "Time Deviation/ms")
  ) +
  labs(x     = "Pulse Width/ms",
       title = "Stimulation Temporal Accuracy across Pulse Width")
```

## Discussion

the replications here closely matched the patterns reported in the original figures. the in vitro temp curves (fig 3e) showed the same voltage-dependent heating pattern, with all 4 voltage conditions reaching distinct steady-state temps that were confirmed to be significantly different by ANOVA + tukey (all p < 0.001). the in vivo data (fig 3g) replicated the finding that subcutaneous temps remain well below dangerous levels during continuous operation. the dual-axis timing plot (fig 3p) successfully reproduced the key finding that relative time deviation (%) decreases with longer pulse widths while absolute deviation (ms) stays roughly stable or increases slightly 

one challenge was that fig3p_pct only had 3 replicates in the data vs 4 for fig3p_ms - this may explain minor differences in error bar size compared to the original. additionally, the original fig 3e uses SEM ribbons while this replication used min/max range - a small methodological difference worth noting.
